Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-05 WSSSPE3/report/WSSSPE3_review.tex
adding a comment
add Birgit's talk summary

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lightning Talks} \label{sec:lightning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{need to update for WSSSPE3}
\todo{KEN: Should these be bolded?}
\begin{comment}
\note{
\href{http://wssspe.researchcomputing.org.uk/wssspe3/agenda/}{Slides.}}
\end{comment}

@@

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Letícia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}

@@

\input{wg_appendix_best_practices}
\input{wg_appendix_funding_spec_expert}
%\input{wg_appendix_catalogs}
\input{wg_appendix_industry_interaction}
%\input{wg_appendix_legacy_SW}
\input{wg_appendix_eng_design}
\input{wg_appendix_metrics}
\input{wg_appendix_training}
\input{wg_appendix_software_credit}
\input{wg_appendix_publishing_SW}
\input{wg_appendix_user_community}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lightning Talks} \label{sec:lightning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{need to update for WSSSPE3}
\todo{KEN: Should these be bolded? - dsk: yes, titles can be bolded, but contents under each should not}
\begin{comment}
\note{
\href{http://wssspe.researchcomputing.org.uk/wssspe3/agenda/}{Slides.}}
\end{comment}

@@

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Letícia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}
The concept of sustainability has become a topic of interest in the field of computing, which is
evidenced by the increase in the number of events that focus on the topic - nevertheless, it isn’t
well understood yet. Penzenstadler argued that we often define sustainability too narrowly. Instead,
sustainability at its heart is a systemic concept and must be viewed from a range of different
dimensions including environmental, economic, individual, social and technical. She introduced the
Karlskrona Manifesto on Software Design, which distills knowledge from a broad range of related work
on the topic of sustainability into a set of (mis-)perceptions and principles. The manifesto does
not proclaim that there is an easy, one size fits all solution around the corner, but rather points
out that sustainability is a ``wicked problem'' and is often misunderstood.
Due to these misperceptions, even though sustainability's importance is increasingly recognized,
many software systems are unsustainable. Even more alarming is that most software systems'
broader impacts on sustainability are unknown. To change this, the Karlskrona Manifesto proposes
nine principles and commitments. These commitments are not dogmatic laws, but rather a commitment to
rethink, to move beyond the silo mentality, and analyze in more depth. As such, they do not
restrict, but rather open up a space for discussion.

@@

\input{wg_appendix_best_practices}
\input{wg_appendix_funding_spec_expert}
%\input{wg_appendix_catalogs}
\input{wg_appendix_industry_interaction}
%\input{wg_appendix_legacy_SW}
\input{wg_appendix_eng_design}
%\input{wg_appendix_metrics}
\input{wg_appendix_training}
\input{wg_appendix_software_credit}
\input{wg_appendix_publishing_SW}
\input{wg_appendix_user_community}
